---
title: "Cross-check data"
author: "Max Lindmark & Federico Mailoi"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 12,
  fig.asp = 0.618,
  fig.align ='center'
)
```
```{r,include=FALSE}
library(tidyverse)
```


## Read data

```{r}
dat=read.csv2("../data/benthic_data_complete.csv",sep=',',row.names = 1)
```

## Wide format

The wide format is convenient for Hmsc, GLLVM, and Vegan although is not *tidy*.
To have a proper wide format we first need to have all abundances/biomass. Either we impute back and forth from abu to biomass by taking the average per taxonomic group, we remove missing data,  or we just use presence-absence data for now.  
We create a list of dataframes. [[1]] will have the long format. [[2]] will be presence-absence data in the wide format. [[3]] abundance wide format. [[4]] biomass wide format.

```{r}
data=list()
data[[1]]=dat
```

We first create the presence absence data wide

```{r}
# any duplicates?
dat %>%
    dplyr::group_by(year, month, day, lon, lat, depth, prov_nr, provID, SubDiv, species_group) %>%
    dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
    dplyr::filter(n > 1L) 
# yep, compress data via distinct()

```

```{r}
dat$pres=1
dat.wide.pa=tidyr::pivot_wider(dat, id_cols = c('year','month','day','lon','lat','depth','prov_nr','provID','SubDiv'), names_from = 'taxa',values_from = 'pres',values_fill = 0)
data[[2]]=dat.wide.pa
```

## Check species prevalence (frequency of occurrence)
```{r}
par(mar = c(12, 5, 5,5)) 
#colnames(data[[2]])
prevalence=(colSums(data[[2]][,-c(1:9)])/dim(data[[2]])[1])*100
barplot(rev(sort(prevalence)),las=2)
abline(h=1,col='red') # line at 1 %
```





